One-The-Fly Decision Procedures For GKAT

### 2024-12-17 main.tex
update

\newcommand\altxrightarrow[2][0pt]{\mathrel{\ensurestackMath{\stackengine%
  {\dimexpr#1-7.5pt}{\xrightarrow{\phantom{#2}}}{\scriptstyle\!#2\,}%
  {O}{c}{F}{F}{S}}}}
\newcommand{\transvia}[1]{
    \mathrel{\raisebox{-2px}{\(\altxrightarrow[-2px]{#1}\)}}
}

@@

@@

\begin{proof}
    Because homomorphic image preserves principle sub-GKAT coalgebra~\cref{thm:homo-img-preserve-principle-sub-coalg}
    \[h(⟨s⟩_S) = ⟨h(s)⟩_T;\]
    therefore for any state \(s' ∈ S\):
    \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_T.\]
    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(h(⟨s⟩_S) = ⟨h(s)⟩_T\) contains an accepting state. 
    Therefore \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(T\).
\end{proof}

@@

    The condition~\labelcref{itm:transition-bisim} holds, by case analysis on the liveness of \(s'\) and \(t'\).
    First note that \(s'\) and \(t'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\), then \(\norm(δ_S)(s', α)\) can either be rejection or \((s',p)\), and so is \(\norm(δ_T)(t', α)\). Finally because \(s ∼ t\), then
    \[s' \text{ is live} ⟺ \norm(δ_S)(s, α) = (s', p) ⟺ \norm(δ_T)(t, α) = (t', p) ⟺ t' \text{ is live}.\]
    \begin{itemize}
        \item If both \(s'\) and \(t'\) are live, then \(s' ∼ t'\), the bisimulation \(∼_{s', t'}\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨t'⟩\).
        \item If both \(s'\) and \(t'\) are dead, then \(∼_{s', t'}\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
    \end{itemize}

@@

Given the alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) consists of a state set \(S\) and a accepting function \(ϵ̂\) and a transition function \(δ̂\):
\begin{mathpar}
    ϵ̂: S → \Bool_B, \and
    δ̂: S → \Bool_B × S × K,
\end{mathpar}
where \(\Bool_B\) is the free boolean algebra over \(B\) (boolean expressions modulo boolean algebra axioms); for all states \(s ∈ S\), all the booleans are ``disjoint''; namely the conjunction of any two expression from the set \(\{ϵ̂(s)\} ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are false. 
We will then use \(ρ̂(s): \Bool_B\) to denote the boolean expressions that contain all the atoms that the state \(s\) rejects, and \(ρ̂(s)\) can be computed as follows:
\[ρ̂(s) ≜ ¬ ϵ̂(s) ∨ ¬ \left( ⋁_{(b, s', p) ∈ δ(s)} b \right)\]

@@

With the above intuition in mind, a symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) can be lowered into a GKAT coalgebra \(⟨S, δ⟩\) in the following manner:
\[
δ(s, α) ≜ \begin{cases}
    \accept & α ≤ ϵ̂(s) \\  
    (s', p) & ∃ b ∈ \Bool_B, α ≤ b \text{ and } δ(s, b) = (s', p) \\  
    \reject & \text{otherwise}
\end{cases}
\]
This is well-defined, i.e. no more than one clause can be satisfied precisely because the boolean expressions appear in \(ϵ̂\) and \(δ̂\) are disjoint.
The trace semantics of a GKAT coalgebra \(⟨S, ϵ̂, δ̂⟩\) is then defined as the trace semantics of its lowering \(⟨S, δ⟩\).

@@

\begin{theorem}[Symbolic Inductive Construction]\label{thm:symb-inductive-construction}
    Given two symbolic GKAT coalgebra \(Ŝ = ⟨S, ϵ̂_S, δ̂_S⟩\) and \(T̂ = ⟨T, ϵ̂_T, δ̂_T⟩\) and two states \(s ∈ S\) and \(t ∈ T\), there exists a normalized bisimulation on the lowered coalgebra \({∼} ⊆ S × T\) s.t. \(s ∼ t\) if and only if all the following holds:
    \begin{itemize}
        \item \(ϵ̂_S(s) ≡ ϵ̂_T(t)\);
        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p = q\) then here exists a normalized bisimulation \({∼_{s',t'}} ⊆ S × T\) s.t. \(s' ∼_{s',t'} t'\);
        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p ≠ q\) then both \(s'\) and \(t'\) is dead;  
        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \(c ∈ ρ̂_T(t)\), if \(b ∧ c ≢ 0\), then \(s'\) is dead;
        \item for all \(b ∈ ρ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\), then \(t'\) is dead;
    \end{itemize}
\end{theorem}

@@

\begin{algorithm}
    \caption{On-the-fly bisimulation algorithm}\label{alg:bisim}
    \begin{algorithmic}
        \Function{equiv}{$s ∈ S, t ∈ T$}
        \If {\Call{eq}{$s, t$}} {\Return true}
        \ElsIf {\(s ∈\) deadStates\(_S\)} {\Return \Call{isDead\(_T\)}{$t$}} 
        \ElsIf {\(t ∈\) deadStates\(_T\)} {\Return \Call{isDead\(_S\)}{$s$}} 
        \Else {}
        \Return {
            \Comment{conditions of ~\cref{thm:symb-inductive-construction}}
            \\\vspace{5px}
            \(\qquad
            \begin{aligned}
                & ϵ̂_S(s) ≡ ϵ̂_T(t) \mathrel{\&\!\&} \\  
                & \forall (b, s', p) ∈ δ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ 
                \begin{cases}
                    \text{\Call{IsDead$_S$}{$s$}} ∧ \text{\Call{IsDead$_T$}{$t$}} & \text{if \(p ≠ q\)} \\
                    \text{\Call{Union}{$s$, $t$}}; \text{\Call{Equiv}{$s', t'$}} & \text{if \(p = q\)}
                \end{cases} \mathrel{\&\!\&}\\
                & \forall (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
                & \forall b ∈ ρ̂_S(s), (c, t', q) ∈ ρ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_T$}{$t'$}}
            \end{aligned}\)
        }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

\subsection{Weak Symbolic GKAT Automata And Tightening}

Weak Symbolic GKAT automata is a symbolic GKAT automata that allow internal transitions via a boolean expression:
\[⟨S, 
ϵ̂: S → \Bool, 
δ̂: S → (\Bool × S) + (\Bool × S × K)⟩.\]
We also require the deterministic criteria, all the boolean transitions of \(s\), i.e. all the boolean expressions in the following set:
\[\{ϵ̂(s)\} ∪ 
\{b ∣ ∃ s' ∈ S, (b, s') ∈ δ̂(s)\} ∪ 
\{b ∣ ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ δ̂(s)\},\]
are disjoint.
Notice that our notion of

>>>>

\newcommand\altxrightarrow[2][0pt]{\mathrel{\ensurestackMath{\stackengine%
  {\dimexpr#1-7.5pt}{\xrightarrow{\phantom{#2}}}{\scriptstyle\!#2\,}%
  {O}{c}{F}{F}{S}}}}
\newcommand{\transvia}[1]{
    \mathrel{\raisebox{-2px}{\(\altxrightarrow[-2px]{#1}\)}}
}
\newcommand{\transAcc}[2]{⇒_{#1} #2}

@@

\begin{theorem}
    Given a functionr \(F\) on set, two \(F\)-colagebra \(S\) and \(T\), if there exists a surjective homomorphism \(h: S → T\), then \(S\) and \(T\) are bisimular.
\end{theorem}

@@

\begin{proof}
    Because homomorphic image preserves principle sub-GKAT coalgebra (\cref{thm:homo-img-preserve-principle-sub-coalg})
    \[h(⟨s⟩_S) = ⟨h(s)⟩_T;\]
    therefore for any state \(s' ∈ S\):
    \[s' ∈ ⟨s⟩_S ⟺ h(s') ∈ h(⟨s⟩_S) ⟺ h(s') ∈ ⟨h(s)⟩_T.\]
    And because \(s'\) is accepting if and only if \(h(s')\) accepting by definition of homomorphism; then \(⟨s⟩_S\) contains an accepting state if and only if \(h(⟨s⟩_S) = ⟨h(s)⟩_T\) contains an accepting state. 
    Therefore \(s\) is live in \(S\) if and only if \(h(s)\) is live in \(T\).
\end{proof}

@@

    The condition~\labelcref{itm:transition-bisim} holds, by case analysis on the liveness of \(s'\) and \(t'\).
    First note that \(s'\) and \(t'\) has to be both live or both dead: because \(δ_S(s, α) = (s', p)\), then \(\norm(δ_S)(s', α)\) can either be rejection or \((s',p)\), and so is \(\norm(δ_T)(t', α)\). Finally because \(s ∼ t\), then
    \[s' \text{ is live} ⟺ \norm(δ_S)(s, α) = (s', p) ⟺ \norm(δ_T)(t, α) = (t', p) ⟺ t' \text{ is live}.\]
    \begin{itemize}
        \item If both \(s'\) and \(t'\) are live, then \(s' ∼ t'\). By~\cref{thm:sub-coalg-preserve-bisim}, the bisimulation \(∼_{s', t'}\) is just \(∼\) restricted to \(⟨s'⟩\) and \(⟨t'⟩\).
        \item If both \(s'\) and \(t'\) are dead, then \(∼_{s', t'}\) can just be the singleton relation, according to~\cref{thm:bisim-between-dead}.
    \end{itemize}

@@

Given the alphabet \(K, B\), a \emph{symbolic GKAT coalgebra} \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) consists of a state set \(S\) and a accepting function \(ϵ̂\) and a transition function \(δ̂\):
\begin{mathpar}
    ϵ̂: S → 𝒫(\Bool_B), \and
    δ̂: S → 𝒫(\Bool_B × S × K),
\end{mathpar}
where \(\Bool_B\) is the free boolean algebra over \(B\) (boolean expressions modulo boolean algebra axioms); for all states \(s ∈ S\), all the booleans are ``disjoint''; namely the conjunction of any two expression from the set \(\{ϵ̂(s)\} ∪ \{b ∣ ∃ (b, s', p) ∈ δ(s)\}\) are false. 
We will then use \(ρ̂(s): \Bool_B\) to denote the boolean expressions that contain all the atoms that the state \(s\) rejects, and \(ρ̂(s)\) can be computed as follows:
\[ρ̂(s) ≜ ¬ ϵ̂(s) ∨ ¬ \left( ⋁_{(b, s', p) ∈ δ(s)} b \right)\]

@@

With the above intuition in mind, a symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) can be lowered into a GKAT coalgebra \(⟨S, δ⟩\) in the following manner:
\[
δ(s, α) ≜ \begin{cases}
    \accept & ∃ b ∈ ϵ̂(s), α ≤ b \\  
    (s', p) & ∃ b ∈ \Bool_B, α ≤ b \text{ and } δ(s, b) = (s', p) \\  
    \reject & \text{otherwise}
\end{cases}
\]
This is well-defined, i.e. no more than one clause can be satisfied precisely because the boolean expressions appear in \(ϵ̂\) and \(δ̂\) are disjoint.
The trace semantics of a GKAT coalgebra \(⟨S, ϵ̂, δ̂⟩\) is then defined as the trace semantics of its lowering \(⟨S, δ⟩\).

@@

\begin{theorem}[Symbolic Inductive Construction]\label{thm:symb-inductive-construction}
    Given two symbolic GKAT coalgebra \(Ŝ = ⟨S, ϵ̂_S, δ̂_S⟩\) and \(T̂ = ⟨T, ϵ̂_T, δ̂_T⟩\) and two states \(s ∈ S\) and \(t ∈ T\), there exists a normalized bisimulation on the lowered coalgebra \({∼} ⊆ S × T\) s.t. \(s ∼ t\) if and only if all the following holds:
    \begin{itemize}
        \item \(⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_T(t)\);
        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p = q\) then here exists a normalized bisimulation \({∼_{s',t'}} ⊆ S × T\) s.t. \(s' ∼_{s',t'} t'\);
        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\) and \(p ≠ q\) then both \(s'\) and \(t'\) is dead;  
        \item for all \((b, s', p) ∈ δ̂_S(s)\) and \(c ∈ ρ̂_T(t)\), if \(b ∧ c ≢ 0\), then \(s'\) is dead;
        \item for all \(b ∈ ρ̂_S(s)\) and \((c, t', q) ∈ δ̂_T(t)\), if \(b ∧ c ≢ 0\), then \(t'\) is dead;
    \end{itemize}
\end{theorem}

@@

\begin{algorithm}
    \caption{On-the-fly bisimulation algorithm}\label{alg:bisim}
    \begin{algorithmic}
        \Function{equiv}{$s ∈ S, t ∈ T$}
        \If {\Call{eq}{$s, t$}} {\Return true}
        \ElsIf {\(s ∈\) deadStates\(_S\)} {\Return \Call{isDead\(_T\)}{$t$}} 
        \ElsIf {\(t ∈\) deadStates\(_T\)} {\Return \Call{isDead\(_S\)}{$s$}} 
        \Else {}
        \Return {
            \Comment{conditions of ~\cref{thm:symb-inductive-construction}}
            \\\vspace{5px}
            \(\qquad
            \begin{aligned}
                & ⋁ ϵ̂_S(s) ≡ ⋁ ϵ̂_T(t) \mathrel{\&\!\&} \\  
                & \forall (b, s', p) ∈ δ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ 
                \begin{cases}
                    \text{\Call{IsDead$_S$}{$s$}} ∧ \text{\Call{IsDead$_T$}{$t$}} & \text{if \(p ≠ q\)} \\
                    \text{\Call{Union}{$s$, $t$}}; \text{\Call{Equiv}{$s', t'$}} & \text{if \(p = q\)}
                \end{cases} \mathrel{\&\!\&}\\
                & \forall (b, s', p) ∈ δ̂_S(s), c ∈ ρ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_S$}{$s'$}} \mathrel{\&\!\&}\\
                & \forall b ∈ ρ̂_S(s), (c, t', q) ∈ δ̂_T(t), (b ∧ c) ≢ ⊥ ⟹ \text{\Call{IsDead$_T$}{$t'$}}
            \end{aligned}\)
        }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

\subsection{Weak Symbolic GKAT Automata, Star System, and Tightening}

Weak Symbolic GKAT automata is a symbolic GKAT automata that allow internal transitions via a boolean expression:
\[⟨S, 
ϵ̂: S → 𝒫(\Bool), 
δ̂: S → 𝒫(\Bool × S + \Bool × S × K)⟩.\]
We also require the deterministic criteria, all the boolean transitions of \(s\), i.e. all the boolean expressions in the following set:
\[ϵ̂(s) ∪ 
\{b ∣ ∃ s' ∈ S, (b, s') ∈ δ̂(s)\} ∪ 
\{b ∣ ∃ s' ∈ S, p ∈ K, (b, s', p) ∈ δ̂(s)\},\]
are disjoint.
To ease the notation, in a symbolic coalgebra \(Ŝ\) we will write 
\begin{itemize}[nosep]
    \item \(s \transAcc{S}{b}\) to mean that \(b ∈ ϵ̂(s)\).
    \item \(s \transvia{b}_{Ŝ} s'\) to mean that \((b, s') ∈ δ̂(s)\);  
    \item and \(s \transvia{b ∣ p}_{Ŝ} s'\) to mean that \((b, s', p) ∈ δ̂(s)\).
\end{itemize}
We will call transitions (to states) that do not emit actions \emph{unproductive}, and similarly, ones that emit actions will be named \emph{productive}.
Specifically, transitions of the form \(s \transvia{b}_{Ŝ} s'\) are unproductive, and \(s \transvia{b ∣ p}_{Ŝ} s'\) represents productive transitions.
Notice that our notion of empty transitions are different from previous work~\cite{silva_CoalgebraicViewETransitions_2013a,bonchi_KillingEpsilonsDagger_2015a,sokolova_WeakBisimulationActionType_2005}, the empty transitions, i.e. \(δ̂(s) ∈ \Bool × S\), although do not execute any action, but will require a boolean formula to transition.
Furthermore, the boolean formula has a natural meet semi-lattice structure given by the logical and operator.
In this section, we will design a specific ``unproductive transition removal'' procedure.

% \begin{definition}[Star Systems]
%     Given a weak symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) a star system \(Ŝ^* ≜ ⟨S, ϵ̂^*, δ̂^*⟩\) is a system that contains exactly the transitions generated by the following rules:
%     \begin{mathpar}
%         \inferrule[Acc]
%         {s \transAcc{Ŝ}{b}}
%         {s \transAcc{Ŝ^*}{b}}
%         \and
%         \inferrule[StarAcc]
%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transAcc{Ŝ^*}{c}}
%         {s₁ \transAcc{Ŝ^*}{b ∧ c}}
%         \and
%         \inferrule[Unprod]
%         {s₁ \transvia{b}_{Ŝ} s₂}
%         {s₁ \transvia{b}_{Ŝ^*} s₂}
%         \and 
%         \inferrule[StarUnprod]
%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{d}_{Ŝ^*} s₃}
%         {s₁ \transvia{b ∧ d}_{Ŝ^*} s₂}
%         \and 
%         \inferrule[Prod]
%         {s₁ \transvia{b ∣ p}_{Ŝ} s₂}
%         {s₁ \transvia{b ∣ p}_{Ŝ^*} s₂}
%         \and 
%         \inferrule[StarProd]
%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{c ∣ p}_{Ŝ^*} s₃}
%         {s₁ \transvia{b ∧ c ∣ p}_{Ŝ^*} s₂}
%     \end{mathpar}

%     The objects in \emph{the category of star system} are weak symbolic GKAT coalgebra and the morphisms are homomorphisms between star systems.
% \end{definition}

% Intuitively, star systems contain \emph{all} the possible transitions both productive and unproductive from a state, until an action is reached.
% Notice that a star system will not necessarily be a weak symbolic GKAT coalgebra, because it will no longer be deterministic.
% One exception is symbolic GKAT coalgebra, i.e.also weak symbolic systems that doesn't have unproductive transitions, where their star system will just be themselves, because only the \textsc{Acc} and \textsc{Prod} rules can be applied.
% Hence the homomorphisms between symbolic GKAT coalgebra in the category of star systems are exactly the standard coalgebra homomorphism of symbolic GKAT coalgebra.

% \begin{theorem}[Universal Property]
%     Take two  GKAT coalgebra \(Ŝ\) and \(Û\), then 
% \end{theorem}

% \begin{proof}
%     Assume \(s \transAcc{Ŝ^*}{b}\), then 
% \end{proof}

% \begin{definition}[Tightening]
%     For a weak symbolic GKAT coalgebra \(Ŝ\), its tightening \(Ŝ\!↓\) is \(Ŝ^*\) with all the unproductive transitions removed. Formally let \(Ŝ^* ≜ ⟨S, ϵ̂^*, δ̂^*⟩\), then \(Ŝ\!↓ ≜ ⟨S, ϵ̂\!↓, δ̂\!↓⟩\) is defined as follows:
%     \begin{mathpar}
%         ϵ̂\!↓(s) ≜ ϵ̂^*(s) \and
%         δ̂\!↓(s) ≜ \{(b, s', p) ∣ (b, s', p) ∈ δ̂^*(s)\} 
%     \end{mathpar}
% \end{definition}

% \begin{lemma}
%     The tightening \(Ŝ\!↓\) of \(Ŝ\) contains exactly the transitions formed by appling \textsc{Acc}, \textsc{StarAcc}, \textsc{Prod}, and \textsc{StarProd} rules to transitions of \(Ŝ\).
% \end{lemma}

% \begin{proof}
%     First all the transitions formed by the four rules will not contain non-productive transition, thus will be contained in the tightening.

%     Second, consider the derivation tree for a productive transition in \(Ŝ^*\) like \(s₁ \transvia{b ∣ p}_{Ŝ^*} s₂\), the \textsc{Unprod} and \textsc{StarUnprod} cannot be the last derivation, since they do not produce a productive transition; and their result cannot be used by any other rules. Thus, \textsc{Unprod} and \textsc{StarUnprod} rule cannot be applied to derive a productive transition.

%     Hence, all the transitions of \(Ŝ^*\), i.e. all the productive transitions of \(Ŝ^*\), are generated exactly by \textsc{Acc}, \textsc{StarAcc}, \textsc{Prod}, and \textsc{StarProd} rules.
% \end{proof}

% \begin{theorem}
%     For any weak symbolic GKAT coalgebra \(Ŝ\), its tightening \(Ŝ\!↓\) is a symbolic GKAT coalgebra.
% \end{theorem}

% \begin{proof}
%     By definition, the tightened system \(Ŝ\!↓\) only contains productive transition. Thus, we only need to verify the disjointness condition.

% We define the \emph{category of star systems} where the objects are weak symbolic GKAT coalgebras, and the homomorphisms between them are standard coalgebra homomorphism between their star systems.

% \begin{theorem}[Universal Property]\label{thm:tight-universal-prop}
%     For all weak symbolic GKAT coalgebra \(Ŝ\), and symbolic GKAT coalgebra \(Û\!↓\), following universal property holds in the category of star system:
%     \[
%     \begin{tikzcd}[sep=1.25cm]
%         Û\!↓ \ar[dashed,swap]{d}{f} \ar{dr}{f}& \\  
%         Ŝ\!↓ \ar[swap]{r}{γ_{Ŝ}} & Ŝ
%     \end{tikzcd}
%     \]
%     where \(γ_{Ŝ}\) maps each state in \(Ŝ\) to itself.
% \end{theorem}

% \begin{proof}
%     We will need to prove three items: first \(γ_{Ŝ}\) is a star system homomorphism, then \(h\) exists and is unique.

%     By definition, tightening contains all the productive and accepting transitions of the star system,
%     therefore if \(s \transAcc{Ŝ\!↓}{b}\) then \(s \transAcc{Ŝ^*}{b}\);
%     similarly, \(s₁ \transvia{b ∣ p}_{Ŝ\!↓} s₂\) implies \(s₁ \transvia{b ∣ p}_{Ŝ^*} s₂\).
%     Thus, \(γ_{Ŝ}\), which maps each state to itself, is a coalgebra homomorphism the star systems.

%     Because \(γ_{Ŝ}\) maps every state to itself, thus \(f\) is the unique function that can make the diagram commute. 
%     We just need to verify that \(f: Û\!↓ → Ŝ\!↓\) satisfy the rules of a star-system homomorphism.
%     \begin{align*}
%         & u₁ \transAcc{Û\!↓}{b} 
%         ⟹ u₁ \transAcc{Û^*}{b} 
%         ⟹ f(u₁) \transAcc{Ŝ^*}{b}
%         ⟹ f(u₁) \transAcc{Ŝ\!↓}{b} \\   
%         & u₁ \transvia{b ∣ p}_{Û\!↓} u₂
%         ⟹ u₁ \transvia{b ∣ p}_{Û^*} u₂
%         ⟹ f(u₁) \transvia{b ∣ p}_{Ŝ^*} u₂
%         ⟹ f(u₁) \transvia{b ∣ p}_{Ŝ\!↓} u₂
%     \end{align*}
% \end{proof}

% Finally, because the star system of a symbolic GKAT coalgebra is just itself, then the coalgebra homomorphisms between them are all symbolic GKAT coalgebra homomorphisms, therefore we obtain the following adjunction theorem.

% \begin{corollary}
%     The tightening \((-)\!↓\) is a functor and is right adjoint to the embedding from symbolic GKAT coalgebra to star systems.
% \end{corollary}

% \begin{proof}
%     This is mostly a standard result of category theory, but for the sake of completeness, we will recreate the result.
%     Given a weak symbolic GKAT coalgebra homomorphism \(f: Û → Ŝ\), it is also a homomorphism of their star systems by definition, then \(f\!↓: Û\!↓ → Ŝ\!↓\) is the unique arrow that satisfy the following diagram:
%     \[
%     \begin{tikzcd}[sep=1cm]
%         Û\!↓ \ar{r}{γ_U} \ar[dashed,swap]{d}{f\!↓} & Û \ar{d}{f}\\  
%         Ŝ\!↓ \ar{r}{γ_S} & Ŝ 
%     \end{tikzcd}
%     \]
%     Because both \(γ_U\) and \(γ_S\) maps states to itself, therefore \(f\!↓\) is necessarily point-wise equal to \(f\); and the preservation of identity and composition can be verified by the universal property.
%     Finally, the adjunction is a concequence of~\Cref{thm:tight-universal-prop}.
% \end{proof}

% The functoriality of tightening is particularly important as it not only means all the homomorphism between the weak coalgebra is also a homomorphism between their tightened version, but also means that it will preserve injectivity and surjectivity as these maps are functions between set. 
% In the next section, we will prove the derivatives are more ``efficient'' than Thomposon's construction by proving there exists a surjective homomorphism from the automata created by derivarives to the automata created by Thompson's construction.
% Such a homomorphism will also be a witness of the bisimularity of the derivatives and Thompson's construction, aiding us in the proof of correctness.

\begin{figure*}
    \begin{mathpar}
        \inferrule[Acc]
        {s₁ \transAcc{Ŝ}{b}}
        {s₁ \transAcc{Ŝ↓}{b}}
        \and
        \inferrule[TightenAcc]
        {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transAcc{Ŝ↓}{c}}
        {s₁ s₂ \transAcc{Ŝ↓}{b ∧ c}}
        \and
        \inferrule[Trans]
        {s₁ \transvia{b ∣ p}_{Ŝ} s₂}
        {s₁ \transvia{b ∣ p}_{Ŝ↓} s₂}
        \and
        \inferrule[TightenTrans]
        {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{c ∣ p}_{Ŝ↓} s₃}
        {s₁ \transvia{b ∧ c ∣ p}_{Ŝ↓} s₃}
    \end{mathpar}
    \caption{Tightening Rules}\label{fig:tightening-rules}
\end{figure*}

\begin{definition}
    Given a weak symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\), its tightening \(Ŝ\!↓ ≜ ⟨S, ϵ̂\!↓, δ̂\!↓⟩\) is a symbolic GKAT coalgebra s.t. \(ϵ̂\!↓\) and \(δ̂\!↓\) are the point-wise least (ordered by set inclusion) function that satisfy the rules in~\cref{fig:tightening-rules} for all \(s₁, s₂, s₃ ∈ S\),
\end{definition}

In particular, infinite loop of unproductive transition will be resolved to rejection, by the definition above.
\begin{example}

\end{example}

One of the helpful observation is that tightening is a functor from the category of weak symbolic GKAT coalgebra into the category of symbolic GKAT colagebra.
The functoriality us to establish isomorphism on the weak case and automatically transfer it to symbolic GKAT coalgebra.

\begin{theorem}[Functoriality]\label{thm:tight-functor-induction}
    Given two weak symbolic GKAT coalgebra \(Ŝ, T̂\) and a homomorphism between them \(h: Ŝ → T̂\), then \(h\) is also a homomorphism on the tightening \(h: Ŝ\!↓ → T̂\!↓\).
\end{theorem}

\begin{proof}
    We will need to show that \(h\) preserves both \(ϵ̂\) and \(δ̂\!↓\). 
    Since the tightening do not change \(ϵ̂\), thus the perseverance of \(ϵ̂\) is trivial.
    Then \(h\) preserves \(δ̂\!↓\) can be shown in two case: consider a transition \(s \transvia{b ∣ p}_{Ŝ\!↓} s'\),
    then there exists a path from \(s\) to \(s'\) in \(Ŝ\) which produced this transition.
    Thus, we can prove the perseverance by induction on the length of the path.
\end{proof}

% \subsection{An Aside, The Categorical Structure of Tightening}

% In this section we will explore the categorical structure of tightening, where we will show the categorical structure of tightening. 
% In particular, we will not only show that \((-)\!↓\) is a functor, but a right adjoint in the appropriate category, without using induction on path as in~\cref{thm:tight-functor-induction}.
% To exhibit the connection between tightened coalgebra and the weak coalgebra, we will need a weaker version of homomorphism. 
% However, before that, we will need to define an alternative way to manipulate weak symbolic colagebra

% \begin{definition}[Star Systems]
%     Given a weak symbolic GKAT coalgebra \(Ŝ ≜ ⟨S, ϵ̂, δ̂⟩\) a star system \(Ŝ^* ≜ ⟨S, ϵ̂^*, δ̂^*⟩\) is a system that contains exactly the transitions generated by the following rules:
%     \begin{mathpar}
%         \inferrule[Acc]
%         {s \transAcc{Ŝ}{b}}
%         {s \transAcc{Ŝ^*}{b}}
%         \and
%         \inferrule[StarAcc]
%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transAcc{Ŝ^*}{b}}
%         {s₁ \transAcc{Ŝ^*}{b}}
%         \and
%         \inferrule[Unprod]
%         {s₁ \transvia{b}_{Ŝ} s₂}
%         {s₁ \transvia{b}_{Ŝ^*} s₂}
%         \and 
%         \inferrule[StarUnprod]
%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{d}_{Ŝ^*} s₃}
%         {s₁ \transvia{b ∧ d}_{Ŝ^*} s₂}
%         \and 
%         \inferrule[Prod]
%         {s₁ \transvia{b ∣ p}_{Ŝ} s₂}
%         {s₁ \transvia{b ∣ p}_{Ŝ^*} s₂}
%         \and 
%         \inferrule[StarProd]
%         {s₁ \transvia{b}_{Ŝ} s₂ \\ s₂ \transvia{c ∣ p}_{Ŝ^*} s₃}
%         {s₁ \transvia{b ∧ c ∣ p}_{Ŝ^*} s₂}
%     \end{mathpar}
% \end{definition}

% Intuitively, star systems contain \emph{all} the possible transitions both productive and unproductive from a state, until an action is reached.
% Notice that a star system will not necessarily be a weak symbolic GKAT coalgebra, because it will no longer be deterministic.
% One exception to these are symbolic GKAT coalgebra, i.e. weak symbolic systems that doesn't have unproductive transitions, where their star system will just be themselves, because only the \textsc{Acc} and \textsc{Prod} rules can be applied.
% Hence the homomorphisms between symbolic GKAT coalgebra in the category of star systems are exactly the standard coalgebra homomorphism of symbolic GKAT coalgebra

% We define the \emph{category of star systems} where the objects are weak symbolic GKAT coalgebras, and the homomorphisms between them are standard coalgebra homomorphism between their star systems.

% \begin{example}
%     some example about \(γ_{Ŝ}\)
% \end{example}

% \begin{theorem}[Universal Property]
%     For all weak symbolic GKAT coalgebra \(Ŝ\), and symbolic GKAT coalgebra \(T̂\!↓\), following universal property holds in the category of star system:
%     \[
%     \begin{tikzcd}[sep=1.25cm]
%         T̂\!↓ \ar[dashed,swap]{d}{f} \ar{dr}{f}& \\  
%         Ŝ\!↓ \ar[swap]{r}{γ_{Ŝ}} & Ŝ
%     \end{tikzcd}
%     \]
%     where \(γ_{Ŝ}\) maps each state in \(Ŝ\) to itself.
% \end{theorem}

% \begin{proof}
%     We will need to prove three items: first \(γ_{Ŝ}\) is a star system homomorphism, then \(h\) exists and is unique.

%     Notice that every the proof rules of tightening is also a proof rule for star system. 
%     Therefore, if \(s \transAcc{Ŝ\!↓}{b}\) can be derived using the rules of tightening, the same trasition can also be derived using the proof rules of the star system.
%     In other word, \(s \transAcc{Ŝ\!↓}{b}\) implies \(s \transAcc{Ŝ^*}{b}\).
%     Similarly \(s₁ \transvia{b ∣ p}_{Ŝ\!↓} s₂\) implies \(s₁ \transvia{b ∣ p}_{Ŝ^*} s₂\).
%     Thus, \(γ_{Ŝ}\), which maps each state to itself, is a morphism in the category of star systems.

%     Because \(γ_{Ŝ}\) maps every state to itself, thus \(f\) is the unique function that can make the diagram commute. 
%     We just need to verify that \(f\) indeed is a star-system homomorphism.

% \begin{corollary}
%     The tightening \((-)\!↓\) is a functor and is right adjoint to the embedding from symbolic GKAT coalgebra to star systems.
% \end{corollary}

\subsection{Derivative, Thomposon's Construction, and Their Relation}

In this section, we will demonstrate the construction of Derivative and Thompson's construction in a symbolic setting.
In addition to the construction itself, we will show there is a surjective homomorphism from Thompson's construction to derivative.
The existence of such surjective homomorphism, especially its lowering by functoriality of lowering, serves several purposes: first, it proves that the semantics of Thompson's construction and derivatives coincides, allowing us to only prove correctness for one of the semantics, instead of both;  
second, it shows that derivative construction produces a smaller automaton than Thompson's construction; finally, it allow us to establish the finiteness of derivatives with the finiteness of Thompson's construction.
